---
title: 'Final Project - Medical Insurance'
author: "Himanshu Chhabra"
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: united
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
---
  
```{r setup, include=FALSE,fig.width=11,echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r Package Block , warning=FALSE,message=FALSE}
library(tidyverse)
library(caret)
library(GGally)
library(readr)
library(tibble)
library(purrr)
library(randomForest)
library(glmnet)
```

## Executive Summary

### Problem and Approach    

Medical Insurance

### Major Results     

Predicting the charges

```{r}
# Loading Dataset
set.seed(3238)
insurance <- read_csv("insurance/insurance_v3.csv")
head(insurance)
```

# 1. Exploratory Data Analysis

## Data Description

```{r results='hide'}
str(insurance)
```
The dataset contains `r nrow(insurance)` observations and `r ncol(insurance)` variables.

## Summary Statistics
```{r}
summary(insurance)
```

## Feature Engineering

```{r}
insurance$sex = as.factor(insurance$sex)
insurance$smoker = as.factor(insurance$smoker)
insurance$region = as.factor(insurance$region)

insurance$bmi30 = as.factor(ifelse(insurance$bmi>30,"Obese","Not Obese"))
```

## Sampling Data : Train/Test Split (80-20) and Normalization

```{r}
sample_index = sample(nrow(insurance), nrow(insurance)*0.8)
insurance_train = insurance[sample_index,] 
insurance_test = insurance[-sample_index,]

x_train = insurance_train %>% dplyr::select(-c(charges))
y_train = insurance_train %>% dplyr::select(charges)

x_test = insurance_test %>% dplyr::select(-c(charges))
y_test = insurance_test %>% dplyr::select(charges)

#Normalize for train data
insurance_train$charges = (insurance_train$charges - min(insurance_train$charges))/(max(insurance_train$charges)-min(insurance_train$charges))
 
 #Normalize for test data
insurance_test$charges = (insurance_test$charges - min(insurance_test$charges))/(max(insurance_test$charges)-min(insurance_test$charges))
```

## Linear Model

```{r}
#Full Model without Interaction
full_model = lm(charges~., data = insurance_train)
#summary(full_model)

#Interaction Effects
formulaWithInterEffects = as.formula(charges ~.+bmi30:smoker)
insurance.interEffect = lm(formulaWithInterEffects, data= insurance_train)
#summary(insurance.interEffect)

#In sample MSE
insurance.lm.pred.train<- predict(insurance.interEffect, insurance_train)
mean((insurance_train$charges-insurance.lm.pred.train)^2)

#Out of Sample MSE
insurance.lm.pred.test<- predict(insurance.interEffect, insurance_test)
mean((insurance_test$charges-insurance.lm.pred.test)^2)
```

## 1. RANDOM FOREST ----------------------------------------------------------------------------

```{r}
oob.err<- rep(0, 7)
test.err<- rep(0, 7)
for(i in 1:7){
  fit<- randomForest(formulaWithInterEffects, data = insurance_train, mtry=i)
  oob.err[i]<- fit$mse[500]
  test.err[i]<- mean((insurance_test$charges-predict(fit, insurance_test))^2)
  cat(i, " ")
}

matplot(cbind(test.err, oob.err), pch=15, col = c("red", "blue"), type = "b", ylab = "MSE", xlab = "mtry")
legend("topright", legend = c("test Error", "OOB Error"), pch = 15, col = c("red", "blue"))

insurance.rf = randomForest(formulaWithInterEffects,data=insurance_train,mtry=2,ntree=200)

plot(insurance.rf$mse,type='l',col=2,lwd=2,xlab="ntree",ylab="OOB Error")

#Variable Importance Plot
varimp <- data.frame(insurance.rf$importance)
ggplot(varimp, aes(x=reorder(rownames(varimp),IncNodePurity), y=IncNodePurity)) +
  geom_bar(stat="identity", fill="blue", colour="black") +
  coord_flip() + theme_bw(base_size = 8) +
  labs(title="Prediction using RandomForest with 200 trees", subtitle="Variable importance (IncNodePurity)", x="Variable", y="Variable importance (IncNodePurity)")

#Predict
insurance.rf.pred.train<- predict(insurance.rf, insurance_train)
insurance.rf.pred.test<- predict(insurance.rf, insurance_test)

# R^2 train
res <- insurance.rf.pred.train - insurance_train$charges
modelOutput_rf.train <- data.frame(obs = insurance_train$charges, pred = insurance.rf.pred.train)


results<-data.frame("RMSE" = 0, "R squared"= 0 , "MAE"= 0)
results<-rbind(results,defaultSummary(modelOutput_rf.train))
results<-results[-1,]

# R^2 test
res <- insurance.rf.pred.test - insurance_test$charges
modelOutput_rf.test <- data.frame(obs = insurance_test$charges, pred = insurance.rf.pred.test)
results<-rbind(results,defaultSummary(modelOutput_rf.test))
```

## 2. XG BOOST ----------------------------------------------------------------------------

```{r}
control<-trainControl(method="cv",number=10)
metric<-"RMSE"

insurance.xgb<-train(formulaWithInterEffects, data =insurance_train, method = "xgbTree",trControl=control,metric=metric, verbose=F) 

insurance.xgb$bestTune

#Predict
insurance.xgb.pred.train<- predict(insurance.xgb, insurance_train)
insurance.xgb.pred.test<- predict(insurance.xgb, insurance_test)

#R2 Train
res <- insurance.xgb.pred.train - insurance_train$charges
modelOutput_xgb.train <- data.frame(obs = insurance_train$charges, pred = insurance.xgb.pred.train)
results<-rbind(results,defaultSummary(modelOutput_xgb.train))


#R2 Test
res <- insurance.xgb.pred.test - insurance_test$charges
modelOutput_xgb.test <- data.frame(obs = insurance_test$charges, pred = insurance.xgb.pred.test)
results<-rbind(results,defaultSummary(modelOutput_xgb.test))

plot(varImp(insurance.xgb),main="Model Feature Importance-XGB")

# RMSE(predictions, insurance_train$charges)
# getTrainPerf(insurance.xgb)
```
## 3. Gradient Boosting ----------------------------------------------------------------------------

```{r}
insurance.gbm<-train(formulaWithInterEffects, data =insurance_train, method = "gbm",trControl=control,metric=metric,
               verbose=F) 

insurance.gbm$bestTune

#Predict
insurance.gbm.pred.train<- predict(insurance.gbm, insurance_train)
insurance.gbm.pred.test<- predict(insurance.gbm, insurance_test)

#R2 Train
res <- insurance.gbm.pred.train - insurance_train$charges
modelOutput_gbm.train <- data.frame(obs = insurance_train$charges, pred = insurance.gbm.pred.train)
results<-rbind(results,defaultSummary(modelOutput_gbm.train))

#R2 Test
res <- insurance.gbm.pred.test - insurance_test$charges
modelOutput_gbm.test <- data.frame(obs = insurance_test$charges, pred = insurance.gbm.pred.test)
results<-rbind(results,defaultSummary(modelOutput_gbm.test))

#Choose the GBM  model
plot(insurance.gbm,main="GBM")

# RMSE(predictions, insurance_train$charges)
# getTrainPerf(insurance.gbm)

#rf_r2 = R2_Score(insurance_train$charges,insurance.gbm.pred.train)
#rf_r2
```

## 4. SVM : REGRESSOR ----------------------------------------------------------------------------

```{r warning=FALSE}
library(quantmod)
library(e1071)

insurance.svm<- train(formulaWithInterEffects, data =insurance_train, method = "svmRadial",trControl=control,metric=metric, verbose=F) 
insurance.svm$bestTune

#Predict
insurance.svm.pred.train<- predict(insurance.svm, insurance_train)
insurance.svm.pred.test<- predict(insurance.svm, insurance_test)

#R2 Train
res <- insurance.svm.pred.train - insurance_train$charges
modelOutput_svm.train <- data.frame(obs = insurance_train$charges, pred = insurance.svm.pred.train)
results<-rbind(results,defaultSummary(modelOutput_svm.train))

#R2 Test
res <- insurance.svm.pred.test - insurance_test$charges
modelOutput_svm.test <- data.frame(obs = insurance_test$charges, pred = insurance.svm.pred.test)
results<-rbind(results,defaultSummary(modelOutput_svm.test))
#insurance.svm<-svm(formulaWithInterEffects,data=insurance_train,method="linear") 
# perform a grid search 
# (this might take a few seconds, adjust how fine of grid if taking too long)
#tuneResult1 <- tune(svm, formulaWithInterEffects,  data = insurance_train)
# Map tuning results
#plot(tuneResult1)

#insurance.svm<-svm(formulaWithInterEffects,data=insurance_train,method="linear") 
#rf_r2 = R2_Score(insurance_train$charges,insurance.svm.pred.train)
#rf_r2
```

## Model Comparison
```{r}
row.names(results) <- c("RF_Train","RF_Test","XGB_Train","XGB_Test","GBM_Train","GBM_Test","SVM_Train","SVM_Test")

results$model <- c("RF_Train","RF_Test","XGB_Train","XGB_Test","GBM_Train","GBM_Test","SVM_Train","SVM_Test")
results$code <- c("RF_Train","RF_Test","XGB_Train","XGB_Test","GBM_Train","GBM_Test","SVM_Train","SVM_Test")

results$Set_type <- c("Train","Test","Train","Test","Train","Test","Train","Test")

results$Model <- c("RF","RF","XGB","XGB","GBM","GBM","SVM","SVM")

ggplot(data=results , aes( x = Model ,y= R.squared , fill = Set_type)) + geom_bar (stat="identity", position = position_dodge(width = 0.1), colour="black" , width = 0.7) + geom_text(aes( x = Model ,y= R.squared +.03, label = round(R.squared,3))) + ggtitle("Comparison of R-squared for Different Models")
```

```{r}
results
```

